---
title: "p8105_hw5_mk4208"
author: "Meeraj Kothari"
date: "11/6/2019"
output: github_document
---

# Loading Libraries

```{r message=FALSE}
library(tidyverse)
library(viridis)
library(devtools)
library(patchwork)
```

# Problem 1

The code chunk below loads the `iris` dataset from the `tidyverse` package and introduces some missing values in each column. The purpose of this problem is to fill in those missing values.

```{r message=FALSE}
library(tidyverse)

set.seed(10)

iris_with_missing = iris %>% 
  map_df(~replace(.x, sample(1:150, 20), NA)) %>%
  mutate(Species = as.character(Species))
```

The following function that a vector as an argument; replaces missing values using the rules defined above; and returns the resulting vector.

```{r message=FALSE}
fill_missing = function(x) {
  if (is.numeric(x)) {
    replace_na(x, mean(x, na.rm = TRUE))
  } else if (is.character(x)) {
    replace_na(x, "virginica")
  }
}
```

The following code chunk applies the `fill_missing` function to the columns of `iris_with_missing` using a `map` statment. 

```{r message=FALSE}
iris_filled = map_df(iris_with_missing, fill_missing)
```

# Problem 2 

The following code chunk creates a dataframe containing all the file names within the zip folder using the `list.files` function.

```{r message=FALSE}
files = list.files("./data/", pattern = ".csv", full.names = TRUE) 
```

The following code chunk iterates over file names and reads in data for each subject using `purrr::map`, saves the result as a new variable in the dataframe and tidies the result by manipulating files names to include control arm and subject ID

```{r message=FALSE}
files_df = files %>% 
  map_df(read_csv) %>% 
  mutate(filename = tools::file_path_sans_ext(basename(files))) %>%
  separate(filename, into = c("arm", "id"), convert = TRUE) %>% 
  mutate(id = row_number()) %>%
  select(id, arm, everything()) %>%
  pivot_longer(week_1:week_8,
    names_to = "week",
    values_to = "value",
    names_prefix = "week_"
  ) %>% 
  mutate(arm = recode(arm, "con" = "Control", "exp" = "Experimental"))
```

The following code chunk maes a spaghetti plot showing observations on each subject over time. 

```{r}
files_df %>% 
  ggplot(aes(x = week, y = value, color = arm, group = id)) +
  geom_line() + 
  theme_minimal() + 
  xlab("Week") +
  ylab("Observations") +
  ggtitle("Observations on each subject over time") +
  theme(legend.position = "bottom") 
```

Based on the above create plot, we can see that the observations for the experimental arm increase over time on average compared to the observations for the control arm. 

# Problem 3 

```{r}
sim_regression = function(beta1 = 0) {
  
  sim_data = tibble(
    x = rnorm(30, mean = 0, sd = 1),
    y = 2 + beta1 * x + rnorm(30, 0, sd = 50^(1/2))
      )
  
  ls_fit = lm(y ~ x, data = sim_data)
  
  tibble(
    beta1_hat = broom::tidy(ls_fit)[2, 2] %>% as.numeric(),
    p_value = broom::tidy(ls_fit)[2, 5] %>% as.numeric()
  )

}
```


```{r}
sim_results = tibble(beta_1 = c(0, 1, 2, 3, 4, 5, 6)) %>% 
  mutate(output_lists = map(.x = beta_1, ~rerun(10, sim_regression(beta1 = .x))),
         estimate_dfs = map(output_lists, bind_rows)) %>%
  select(-output_lists) %>%
  unnest(estimate_dfs)
```

```{r}
sim_results %>% 
  group_by(beta_1) %>%
  count(reject = p_value < 0.05) %>% 
  mutate(percent = n/sum(n)*100) %>% 
  filter(reject == TRUE) %>%
  ggplot(aes(x = beta_1, y = percent)) +
  geom_point() + 
  geom_line()
```

```{r}
p1 = sim_results %>% 
  group_by(beta_1) %>%
  summarise(mean = mean(beta1_hat)) %>% 
  ggplot(aes(x = beta_1, y = mean)) + 
  geom_point() + 
  geom_line()

p2 = sim_results %>% 
  filter(p_value < 0.05) %>% 
  group_by(beta_1) %>%
  summarise(mean = mean(beta1_hat)) %>% 
  ggplot(aes(x = beta_1, y = mean)) + 
  geom_point() +
  geom_line()

p1 + p2
```

